Astr 513: Final Project

### 2016-05-03 Final_Project/ms.tex
Fixed a few typos and updated the distance modulus equation

\section{Type \rom{1}a Supernovae}
\label{sec:SNIa}
\EW{Should I cite the Nobel Prize information document where I found out about 
the stuff in the last few sentences? I'm not entirely sure what that would be considered...}
\TM{No, because it is common knowledge. I think though that since this is meant to emulate
  a research paper that we don't need to point out that it was what the NP was awarded for.}
A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
helium absorption. It is believed that the progenitor for these special
supernovae are white dwarfs that are part of a binary system. These white
dwarfs accrete mass from their companion stars or merge with a companion white
dwarf or neutron star, causing them to exceed the Chandrasekhar limit
($1.44 M_\odot$). When this happens, carbon and oxygen fusion begins in the
core, and the temperature increases. However, the white dwarf does not expand
and cool as a main sequence star would, since it is supported by degeneracy
pressure rather than thermal, so the core experiences a runaway fusion
reaction, releasing $\sim 10^{44}$ J of energy \citep[see, e.g.,][chap. 18.4]{ryden2010}, 
which is enough to overcome the gravitational energy and the star explodes. Because all 
\snea{} are caused by a similar reaction, they all have similar luminosities, and more 
importantly, the light curves all have the same shape: a sharp maximum followed by a smooth, 
gradual decay. Studying nearby \snea{} has shown that the rate of decline of the light curve 
is correlated to the peak luminosity, and this correlation has been calibrated using Cephied 
variable stars located in the same galaxies as nearby \sna{}. With this information, one can 
estimate peak luminosity of an \sna{} from its light curve, thereby providing a 
standard candle with which to measure distance. They were first used in this way in the 
1990s by two independent teams, the Supernova Cosmology Project and the Cal\'{a}n/Tololo 
Supernova Survey. It was using distances from \snea{} from these two projects that 
\citet{riess1998} and \citet{perlmutter1999} were able to discover the accelerating expansion 
rate of the Universe in 1998. 
%It was for this work that Saul Perlmutter, Brian P. Schmidt, 
%and Adam G. Reiss won the Nobel Prize in physics in 2011.
\TM{Changed grammer in this section, since ``we'' did not actually do these things.
  Review and collapse bubbles.}

@@

The publicly available data from \citet{rest2014} contains all light 
curves, light curve fit information, covariance matrix, and each {\sn} 
host information.\footnote{Data can be accessed at 
\url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}. For our analysis, we 
utilized the redshifts, distance moduli, and covariance matrix of the 
final sample of 310 {{\snea} or 335 light curves.
\TM{There is some kind of undefined reference error in the above line. 
  It doesn't alter the pdf document, but can we figure out what is going on?}

@@

The correlation matrix is calculated from the covariance matrix according to
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
figures \autoref{fig:cmatrix_betoule} and \autoref{cmatrix_rest}.

@@

\section{Distance Modulus Model}
\label{sec:model}
\LC{Need more here but I wrote most of the relevant eqns}
\EW{Do we want/need citations for the equations here?}
The main objective of \citet{betoule2014} and \citet{rest2014} is 
to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
and $w$. In order to so, we must examine the relationship between distance 
modulus and cosmology. The distance modulus $\mu$ is determined from 
the luminosity distance $d_L$ by 
\begin{equation}
\label{eq:mu_dL}
\mu=5 \text{log}(d_L)+25. 
\end{equation}
The luminosity distance depends on the cosmology by 
\begin{equation}
\label{eq:luminosityDistance}
d_L(z; w, \Omega_M, \Omega_\Lambda, H_0)= (1+z)|\Omega_k|^{-1/2}\mathcal{S}_{k}\displaystyle\left[c|\Omega_k|^{1/2}\int_{0}^{z}\frac{dz'}{H(z')}\right],
\end{equation}
where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
$\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
\[
    \mathcal{S}_{k}= 
\begin{cases}
    \text{sin}(x),&  \Omega_k < 0\\
    x,& \Omega_k = 0\\
    \text{sinh}(x),& \Omega_k > 0
\end{cases}
\]
and the function $H(z)$ is 
\begin{equation}
\label{eq:hubbleFunction}
H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2}.
\end{equation}
Three (?) cosmologies are of interest to us: 
\LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
\TM{I think there are only two, one with $\Omega_k=0$ and one with it
  being unconstrained.}

>>>>

\section{Type \rom{1}a Supernovae}
\label{sec:SNIa}
A type \rom{1}a supernova (\sna{}) is a supernova with very little hydrogen or
helium absorption. It is believed that the progenitor for these special
supernovae are white dwarfs that are part of a binary system. These white
dwarfs accrete mass from their companion stars or merge with a companion white
dwarf or neutron star, causing them to exceed the Chandrasekhar limit
($1.44 M_\odot$). When this happens, carbon and oxygen fusion begins in the
core, and the temperature increases. However, the white dwarf does not expand
and cool as a main sequence star would, since it is supported by degeneracy
pressure rather than thermal, so the core experiences a runaway fusion
reaction, releasing $\sim 10^{44}$ J of energy \citep[see, e.g.,][chap. 18.4]{ryden2010}, 
which is enough to overcome the gravitational energy and the star explodes. Because all 
\snea{} are caused by a similar reaction, they all have similar luminosities, and more 
importantly, the light curves all have the same shape: a sharp maximum followed by a smooth, 
gradual decay. Studying nearby \snea{} has shown that the rate of decline of the light curve 
is correlated to the peak luminosity, and this correlation has been calibrated using Cephied 
variable stars located in the same galaxies as nearby \sna{}. With this information, one can 
estimate peak luminosity of an \sna{} from its light curve, thereby providing a 
standard candle with which to measure distance. They were first used in this way in the 
1990s by two independent teams, the Supernova Cosmology Project and the Cal\'{a}n/Tololo 
Supernova Survey. It was using distances from \snea{} from these two projects that 
\citet{riess1998} and \citet{perlmutter1999} were able to discover the accelerating expansion 
rate of the Universe in 1998.

@@

The publicly available data from \citet{rest2014} contains all light 
curves, light curve fit information, covariance matrix, and each {\sn} 
host information.\footnote{Data can be accessed at 
\url{http://ipp0012.ifa.hawaii.edu/SN1aLC/}}. For our analysis, we 
utilized the redshifts, distance moduli, and covariance matrix of the 
final sample of 310 {\snea} or 335 light curves.

@@

The correlation matrix is calculated from the covariance matrix according to
\begin{equation}
  \label{eq:correlation}
  Corr_{i,j} = \frac{\mathbf{C}_{i,j}}{\sqrt{\mathbf{C}_{i,i}}\sqrt{\mathbf{C}_{j,j}}}.
\end{equation}
Note that the correlation of a variable with itself is 1 by definition.
The correlation matrices for both of our data sets can be seen below in
figures \autoref{fig:cmatrix_betoule} and \autoref{fig:cmatrix_rest}.

@@

\section{Distance Modulus Model}
\label{sec:model}
\LC{Need more here but I wrote most of the relevant eqns}
The main objective of \citet{betoule2014} and \citet{rest2014} is 
to place constraints on cosmological parameters $\Omega_M$, $\Omega_\Lambda$, 
and $w$. In order to so, we must examine the relationship between distance 
modulus and cosmology. The distance modulus $\mu$ is determined from 
the luminosity distance $d_L$ by 
\begin{equation}
\label{eq:mu_dL}
\mu=5 \log\left(\frac{d_L}{1 \mathrm{Mpc}}\right)+25. 
\end{equation}
The luminosity distance depends on the cosmology by 
\begin{equation}
\label{eq:luminosityDistance}
d_L(z; w, \Omega_M, \Omega_\Lambda, H_0)= (1+z)|\Omega_k|^{-1/2}\mathcal{S}_{k}\displaystyle\left[c|\Omega_k|^{1/2}\int_{0}^{z}\frac{dz'}{H(z')}\right],
\end{equation}
where $\Omega_k \equiv 1-\Omega_M-\Omega_{\Lambda}$. The function 
$\mathcal{S}_{k}$ depends on the curvature density $\Omega_k$ in that 
\[
    \mathcal{S}_{k}= 
\begin{cases}
    \text{sin}(x),&  \Omega_k < 0\\
    x,& \Omega_k = 0\\
    \text{sinh}(x),& \Omega_k > 0
\end{cases}
\]
and the function $H(z)$ is 
\begin{equation}
\label{eq:hubbleFunction}
H(z)= H_0\displaystyle\left[\Omega_M(1+z)^3+\Omega_{\Lambda}(1+z)^{3(1+w)}+\Omega_k(1+z)^2\right]^{1/2}.
\end{equation}
Two cosmologies are of interest to us: 
\LC{Not sure which ones we're interested in. $\Lambda CDM$, $\Omega_M=1$, ?}
\TM{I think there are only two, one with $\Omega_k=0$ and one with it
  being unconstrained.}
